### 2024-08-08 Codes/0_1_energy_water_merge.rmd
Initial commit. Branching out from the original repository of Energy Regulation Welfare Loss.

---
title: Match energy and water datasets
---

# Preparation

```{r}
#--- set working directory ---#
setwd("~/Dropbox/NebraskaWaterProjectsData")

#--- libraries ---#
source("~/Dropbox/R_libraries_load/library.R")

# ===================================
# Import datasets
# ===================================
#--------------------------
# Well registration
#--------------------------
# Notes:
# When subsec has four letters,
# The first two letters represent subsubseciton
# The last two letters represent subseciton

wells <- fread("./Data/WellRegistrationRaw/Reg08_26_2016_Raw/all_rrb_wells.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  filter(useid == "I") %>% # irrigation wells	only
  select( # keep the relevant variables
    wellid, twnid, rngid, rngdir, section, subsecmnemonic,
    subsecid, nrdname, swl, pwl, totaldepth, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  rename(township = twnid, range = rngid, subsec = subsecmnemonic) %>%
  mutate(
    trs = paste(township, range, section, sep = "_"),
    subsection = ifelse(str_length(subsec) == 2, subsec, str_sub(subsec, 3, 4)),
    trs_s = paste(trs, subsection, sep = "_"),
    subsubsection = ifelse(str_length(subsec) == 2, NA, str_sub(subsec, 1, 2)),
    trs_ss = paste(trs_s, subsubsection, sep = "_")
  ) %>%
  arrange(township, range, section) %>%
  data.table() %>%
  .[, N_sec := .N, by = .(trs)] %>%
  .[, N_sub := .N, by = .(trs_s)] %>%
  .[, N_subsub := .N, by = .(trs_ss)]

# wells[N_sub>4,]
# wells[trs=='1_7_29',]

#++++++++++++++++
# decommission date
#++++++++++++++++
wells <- wells %>%
  mutate(
    dec_date = as.Date(decommissiondate, "%m/%d/%y"),
    dec_year = year(dec_date),
    dec_month = month(dec_date),
    inactive_not_dec = ifelse(is.na(dec_year) & wellstatusid == "X", 1, 0)
  ) %>%
  data.table()

# Notes
# There are wells that are inactive even though they have not been decommissioned
# I probably should get rid of them

#--- visualization of wells per section and subsection ---#
# ggplot(data=wells) +
# 	geom_histogram(aes(x=N_sub),color='blue',fill='white')

#--------------------------
# Water use data
#--------------------------
gw <- readRDS("./Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  .[, extended_id := paste(volaf, acres, year, county)] %>%
  .[, count_id := .N, by = extended_id] %>% # get rid of duplicate records
  .[count_id == 1, ]

# ===================================
# with Midwest
# ===================================
Midwest <- readRDS("./Data/Energy/Processed/Midwest.rds") %>%
  select(year, district, rate, township, range, section, subsection, subsubsection, hp, kwh) %>%
  mutate(
    trs = paste(township, range, section, sep = "_"),
    trs_s = paste(trs, subsection, sep = "_")
  ) %>%
  arrange(township, range, section, subsection, year) %>%
  data.table() %>%
  .[, N_sec := .N, by = .(trs, year)] %>%
  .[, N_sub := .N, by = .(trs_s, year)]

#--- find wells within the Midwest PPD ---#
MW_trs_ls <- Midwest[, trs] %>% unique()
wells_MW <- wells %>%
  filter(trs %in% MW_trs_ls) %>%
  data.table()

#--------------------------
# single trs-trs
#--------------------------
# pick observations that are the only record at the section level (not many)
MW_sec <- Midwest[N_sec == 1, ] %>%
  left_join(., wells_MW[N_sec == 1, ], by = "trs") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs_s = trs_s.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()

# MW_sec[!is.na(wellid),]

#--------------------------
# single trs_s-trs_s
#--------------------------
# pick observations that are the only record at the subsection level
#--- the only energy record at the subsection level ---#
MW_sub <- Midwest[N_sub == 1 & N_sec > 1, ] %>%
  left_join(., wells_MW[N_sub == 1 & N_sec > 1, ], by = "trs_s") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs = trs.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()

# matched_sub <- MW_sub[!is.na(wellid),]
# unmatched_sub <- MW_sub[is.na(wellid),]

#--------------------------
# merge with pumping data
#--------------------------
# only MW_sub and MW_sec for now. later add MW_sub_2
MW_we <- rbind(MW_sec, MW_sub) %>%
  left_join(gw, ., by = c("wellid", "year")) %>%
  filter(!is.na(district)) %>%
  data.table()

#--------------------------
# Matching accuracy check
#--------------------------
MW_we <- MW_we %>%
  .[, w_to_e := volaf * 12 / kwh] %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(wellid)]

well_ls <- MW_we[, wellid] %>% unique()
well_len <- length(well_ls)

#--- visualization ---#
# for (i in 1:well_len){
# 	temp_data <- MW_we[wellid==well_ls[i],]
# 	temp_cv <- temp_data[,coef_var*1e5] %>%
# 		unique() %>%
# 		round(3)
# 	ggplot(data=temp_data) +
# 		geom_line(aes(y=w_to_e,x=year)) +
# 		ylim(0,0.1) +
# 		annotate('text',label=paste('Coef Var = ',temp_cv,sep=''),y=0.075,x=temp_data[,mean(year)])
# 	ggsave(paste('./Graphs/w_to_e/Midwest/consistency_',well_ls[i],'.pdf',sep=''))
# }

# Notes:
# According to visual inspections, coef_var<7e-4 seems acceptable
MW_matched_1st <- MW_we[coef_var < 7e-4, ] %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

#--------------------------
# multiple wells and multiple energy records in the same trs
#--------------------------
matched_well_ls <- MW_matched_1st[, wellid] %>% unique()
matched_trs_s_ls <- MW_matched_1st[, trs_s] %>% unique()

wells_2nd <- wells_MW[!(wellid %in% matched_well_ls), ]
e_2nd <- Midwest[!(trs_s %in% matched_trs_s_ls), ]

temp_match_2nd <- e_2nd[N_sub == 1, ] %>%
  left_join(., wells_2nd, by = "trs") %>%
  left_join(., gw, by = c("wellid", "year")) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  filter(matched == 1) %>%
  rename(trs_s = trs_s.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, volaf, acres, source_txt, source, hp,
    kwh, regdate, pumprate, pumpdepth, decommissiondate, district, rate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  mutate(well_trs_s = paste(wellid, trs_s, sep = "_")) %>%
  mutate(w_to_e = volaf * 12 / kwh) %>%
  filter(!is.na(volaf)) %>%
  arrange(well_trs_s, year) %>%
  data.table() %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(well_trs_s)]

#--- list of the trs ---#
trs_ls <- temp_match_2nd[, trs] %>% unique()
trs_len <- length(trs_ls)

#++++++++++++++++
# loop over trs to find matches
#++++++++++++++++
matched_store_all <- list()

for (i in 1:trs_len) { # loop over trs
  #--- select well-energy records combinations in the temporary trs ---#
  temp_match <- temp_match_2nd[trs == trs_ls[i] & !is.na(coef_var), .(well_trs_s, wellid, trs_s, coef_var)] %>%
    unique(by = "well_trs_s")
  well_len <- temp_match[, wellid] %>%
    unique() %>%
    length()

  matched_store_i <- list()
  for (j in 1:well_len) { # loop over wells
    if (min(temp_match[, coef_var]) < 7e-4) { # if there is a match
      #--- identify the best matching well-energy combination ---#
      min_index <- which.min(temp_match[, coef_var])
      get_well <- temp_match[min_index, wellid]
      get_trs_s <- temp_match[min_index, trs_s]

      #--- extract the matched data and store in the temporary storage ---#
      matched_store_i[[j]] <- temp_match_2nd %>%
        filter(wellid == get_well & trs_s == get_trs_s) %>%
        data.table()

      #--- exclude the matched wellid and energy records  ---#
      temp_match <- temp_match %>%
        filter(!(wellid == get_well | trs_s == get_trs_s)) %>%
        data.table()
    } else { # if no good enough matches
      break # end the loop
    }
  }
  matched_store_all[[i]] <- rbindlist(matched_store_i)
}

MW_matched_2nd <- rbindlist(matched_store_all) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

# ggplot(data=temp_match_2nd[trs==trs_ls[i],]) +
# 	geom_line(aes(y=w_to_e,x=year,color=trs_s)) +
# 	geom_point(aes(y=w_to_e,x=year,color=trs_s)) +
# 	facet_wrap(~wellid,nrow=3) +
# 	ylim(0,0.1) +
# 	theme(
# 		legend.position = 'bottom'
# 		)

#--------------------------
# combine the 1st and 2nd matches and save
#--------------------------
MW_we <- rbind(MW_matched_1st, MW_matched_2nd) %>%
  .[, hours := kwh / (hp * 0.746)] %>%
  .[, gpm := (volaf * 12 * 27154) / (hours * 60)]

saveRDS(MW_we, "./Data/Water_Energy/MW_energy_water.rds")

# ggplot(data=MW_we[w_to_e>0.008 & pwl <500,]) +
# 	geom_point(aes(y=1/w_to_e,x=pumpdepth),size=0.4) +
# 	geom_smooth(aes(y=1/w_to_e,x=pumpdepth))

# ggplot(data=MW_we[w_to_e>0.008 & pwl <500,]) +
# 	geom_point(aes(y=gpm,x=pumprate),size=0.4)

# ggplot(data=MW_we[w_to_e>0.008 & pwl <500,]) +
# 	geom_point(aes(y=gpm,x=hours),size=0.4)

# ggplot(data=MW_we[w_to_e>0.008 & pwl <500,]) +
# 	geom_point(aes(y=1/w_to_e,x=pwl),size=0.4) +
# 	geom_smooth(aes(y=1/w_to_e,x=pwl)) +
# 	xlab('Pumping Water Level (feet)') +
# 	ylab('Energy Use Efficiency (kwh/acre-inch)')

# ggplot(data=MW_we[w_to_e>0.008,]) +
# 	geom_point(aes(y=gpm,x=hp),size=0.4) +
# 	geom_smooth(aes(y=gpm,x=hp))

ggplot(data = MW_we[w_to_e > 0.008, ]) +
  geom_boxplot(aes(y = gpm, x = factor(year)))

# ggplot(data=MW_we[w_to_e>0.008,]) +
# 	geom_density(aes(x=gpm)) +
# 	facet_wrap(~year,nrow=3)

# ===================================
# with McCook
# ===================================
# 2010 data not reliable (too small numbers of kwh)
#--- import energy data ---#
McCook <- readRDS("./Data/Energy/Raw/McCook/Processed/McCook.rds") %>%
  select(year, district, township, range, section, subsection, subsubsection, hp, kwh) %>%
  filter(year != 2010) %>%
  mutate(
    trs = paste(township, range, section, sep = "_"),
    trs_s = paste(trs, subsection, sep = "_"),
    trs_ss = paste(trs_s, subsubsection, sep = "_")
  ) %>%
  arrange(township, range, section, subsection, subsubsection, year) %>%
  data.table() %>%
  .[, N_sec := .N, by = .(trs, year)] %>%
  .[, N_sub := .N, by = .(trs_s, year)] %>%
  .[, N_subsub := .N, by = .(trs_ss, year)]

#--- find wells within the McCook PPD ---#
MC_trs_ls <- McCook[, trs] %>% unique()
wells_MC <- wells %>%
  filter(trs %in% MC_trs_ls) %>%
  data.table()

#--------------------------
# single trs-trs
#--------------------------
# pick observations that are the only record at the section level (not many)
MC_sec <- McCook[N_sec == 1, ] %>%
  left_join(., wells_MC[N_sec == 1, ], by = "trs") %>%
  rename(trs_s = trs_s.x, trs_ss = trs_ss.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year, district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

# MC_sec[matched==1,]

#--------------------------
# single trs_s-trs_s
#--------------------------
# pick observations that are the only record at the subsection level
MC_sub <- McCook[N_sub == 1 & N_sec > 1, ] %>%
  left_join(., wells_MC[N_sub == 1 & N_sec > 1, ], by = "trs_s") %>%
  rename(trs = trs.x, trs_ss = trs_ss.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year, district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

# MC_sub[matched==1,]

#--------------------------
# single trs_ss-trs_ss
#--------------------------
# pick observations that are the only record at the subsubsection level
MC_subsub <- McCook[N_subsub == 1 & N_sub > 1, ] %>%
  left_join(., wells_MC[N_subsub == 1 & N_sub > 1, ], by = "trs_ss") %>%
  rename(trs = trs.x, trs_s = trs_s.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year, district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

# MC_subsub[matched==1,]

#--------------------------
# Merge
#--------------------------
MC_we <- rbind(MC_sec, MC_sub, MC_subsub) %>%
  filter(matched == 1) %>%
  left_join(gw, ., by = c("wellid", "year")) %>%
  filter(!is.na(district)) %>%
  data.table()

#--------------------------
# Matching accuracy check
#--------------------------
MC_we <- MC_we %>%
  .[, w_to_e := volaf * 12 / kwh] %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(wellid)]

# #--- visualization ---#
# well_ls <- MC_we[,wellid] %>% unique()
# well_len <- length(well_ls)
# for (i in 1:well_len){
# 	temp_data <- MC_we[wellid==well_ls[i],]
# 	temp_cv <- temp_data[,coef_var*1e5] %>%
# 		unique() %>%
# 		round(3)
# 	ggplot(data=temp_data) +
# 		geom_line(aes(y=w_to_e,x=year)) +
# 		ylim(0,0.1) +
# 		annotate('text',label=paste('Coef Var = ',temp_cv,sep=''),y=0.075,x=temp_data[,mean(year)])
# 	ggsave(paste('./Graphs/w_to_e/McCook/consistency_',well_ls[i],'.pdf',sep=''))
# }

# Notes:
# According to visual inspections, coef_var<7e-4 seems acceptable
MC_matched_1st <- MC_we[coef_var < 7e-4, ] %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, trs_ss, district, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

#--------------------------
# multiple wells and multiple energy records in the same trs
#--------------------------
matched_well_ls <- MC_matched_1st[, wellid] %>% unique()
matched_trs_ss_ls <- MC_matched_1st[, trs_ss] %>% unique()

wells_2nd <- wells_MC[!(wellid %in% matched_well_ls), ]
e_2nd <- McCook[!(trs_s %in% matched_trs_ss_ls), ]

temp_match_2nd <- e_2nd[N_sub == 1, ] %>%
  left_join(., wells_2nd, by = "trs") %>%
  left_join(., gw, by = c("wellid", "year")) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  filter(matched == 1) %>%
  rename(trs_s = trs_s.x, trs_ss = trs_ss.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year, volaf, acres, source_txt, source, hp,
    kwh, regdate, pumprate, pumpdepth, decommissiondate, district,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  mutate(well_trs_ss = paste(wellid, trs_ss, sep = "_")) %>%
  mutate(w_to_e = volaf * 12 / kwh) %>%
  filter(!is.na(volaf)) %>%
  arrange(well_trs_ss, year) %>%
  data.table() %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(well_trs_ss)]

#--- visual inspection ---#
# i=17
# ggplot(data=temp_match_2nd[trs==trs_ls[i],]) +
# 	geom_line(aes(y=w_to_e,x=year,color=trs_s)) +
# 	geom_point(aes(y=w_to_e,x=year,color=trs_s)) +
# 	facet_wrap(~wellid,nrow=3) +
# 	ylim(0,0.1) +
# 	theme(
# 		legend.position = 'bottom'
# 		)

# head(temp_match_2nd[trs==trs_ls[i],],50)

#++++++++++++++++
# loop over trs to find matches
#++++++++++++++++
#--- list of the trs ---#
trs_ls <- temp_match_2nd[, trs] %>% unique()
trs_len <- length(trs_ls)

matched_store_all <- list()
for (i in 1:trs_len) { # loop over trs
  #--- select well-energy records combinations in the temporary trs ---#
  temp_match <- temp_match_2nd[trs == trs_ls[i] & !is.na(coef_var), .(well_trs_ss, wellid, trs_ss, coef_var)] %>%
    unique(by = "well_trs_ss")
  well_len <- temp_match[, wellid] %>%
    unique() %>%
    length()

  matched_store_i <- list()
  for (j in 1:well_len) { # loop over wells
    if (min(temp_match[, coef_var]) < 7e-4) { # if there is a match
      #--- identify the best matching well-energy combination ---#
      min_index <- which.min(temp_match[, coef_var])
      get_well <- temp_match[min_index, wellid]
      get_trs_ss <- temp_match[min_index, trs_ss]

      #--- extract the matched data and store in the temporary storage ---#
      matched_store_i[[j]] <- temp_match_2nd %>%
        filter(wellid == get_well & trs_ss == get_trs_ss) %>%
        data.table()

      #--- exclude the matched wellid and energy records  ---#
      temp_match <- temp_match %>%
        filter(!(wellid == get_well | trs_ss == get_trs_ss)) %>%
        data.table()
    } else { # if no good enough matches
      break # end the loop
    }
  }
  matched_store_all[[i]] <- rbindlist(matched_store_i)
}

MC_matched_2nd <- rbindlist(matched_store_all) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, trs_ss, district, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table() %>%
  .[, count := .N, by = wellid] %>%
  .[count > 4, ] %>%
  .[, count := NULL]

#--------------------------
# combine the 1st and 2nd matches and save
#--------------------------
MC_we <- rbind(MC_matched_1st, MC_matched_2nd) %>%
  .[, hours := kwh / (hp * 0.746)] %>%
  .[, gpm := (volaf * 12 * 27154) / (hours * 60)] %>%
  .[w_to_e < 0.1 & w_to_e > 0.008, ]

saveRDS(MC_we, "./Data/Water_Energy/MC_energy_water.rds")

# ===================================
# with Southern
# ===================================
Southern <- readRDS("./Data/Energy/Processed/Southern.rds") %>%
  .[, district := "Southern"]

#--- find wells within the Midwest PPD ---#
SO_trs_ls <- Southern[, trs] %>% unique()
wells_SO <- wells %>%
  filter(trs %in% SO_trs_ls) %>%
  data.table()

#--------------------------
# single trs-trs
#--------------------------
# pick observations that are the only record at the section level (not many)
SO_sec <- Southern[N_trs == 1, ] %>%
  left_join(., wells_SO[N_sec == 1, ], by = "trs") %>%
  rename(trs_s = trs_s.x, trs_ss = trs_ss.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year,
    district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

#--------------------------
# single trs_s-trs_s
#--------------------------
# pick observations that are the only record at the subsection level
SO_sub <- Southern[N_trs_s == 1 & N_trs > 1, ] %>%
  left_join(., wells_SO[N_sub == 1 & N_sec > 1, ], by = "trs_s") %>%
  rename(trs = trs.x, trs_ss = trs_ss.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year,
    district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

# MC_sub[matched==1,]

#--------------------------
# single trs_ss-trs_ss
#--------------------------
# pick observations that are the only record at the subsubsection level
SO_subsub <- Southern[N_trs_ss == 1 & N_trs_s > 1, ] %>%
  left_join(., wells_SO[N_subsub == 1 & N_sub > 1, ], by = "trs_ss") %>%
  rename(trs = trs.x, trs_s = trs_s.x) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year,
    district, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  data.table()

SO_subsub[matched == 1, ]

#--------------------------
# Merge
#--------------------------
SO_we <- rbind(SO_sec, SO_sub, SO_subsub) %>%
  filter(matched == 1) %>%
  left_join(gw, ., by = c("wellid", "year")) %>%
  filter(!is.na(district)) %>%
  data.table()

SO_matched_1st <- SO_we %>%
  .[, w_to_e := volaf * 12 / kwh] %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(wellid)] %>%
  .[coef_var < 7e-4, ] %>%
  .[, count := .N, by = wellid] %>%
  .[count > 4, ] %>%
  .[, count := NULL] %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, trs_ss, district, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

#--- visualization ---#
# well_ls <- SO_matched_1st[,wellid] %>% unique()
# well_len <- length(well_ls)
# for (i in 1:well_len){
# 	temp_data <- SO_matched_1st[wellid==well_ls[i],]
# 	temp_cv <- temp_data[,coef_var*1e5] %>%
# 		unique() %>%
# 		round(3)
# 	ggplot(data=temp_data) +
# 		geom_line(aes(y=w_to_e,x=year)) +
# 		ylim(0,0.1) +
# 		annotate('text',label=paste('Coef Var = ',temp_cv,sep=''),y=0.075,x=temp_data[,mean(year)])
# 	ggsave(paste('./Graphs/w_to_e/Southern/consistency_',well_ls[i],'.pdf',sep=''))
# }

#--------------------------
# multiple wells and multiple energy records in the same trs
#--------------------------
matched_well_ls <- SO_matched_1st[, wellid] %>% unique()
matched_trs_ss_ls <- SO_matched_1st[, trs_ss] %>% unique()

wells_2nd <- wells_SO[!(wellid %in% matched_well_ls), ]
e_2nd <- Southern[!(trs_s %in% matched_trs_ss_ls), ]

temp_match_2nd <- e_2nd[N_trs_ss == 1, ] %>%
  left_join(., wells_2nd, by = "trs") %>%
  left_join(., gw, by = c("wellid", "year")) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  filter(matched == 1) %>%
  rename(trs_s = trs_s.x, trs_ss = trs_ss.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, trs_ss, year, volaf, acres, source_txt, source, hp,
    kwh, regdate, pumprate, pumpdepth, decommissiondate, district,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  mutate(well_trs_ss = paste(wellid, trs_ss, sep = "_")) %>%
  mutate(w_to_e = volaf * 12 / kwh) %>%
  filter(!is.na(volaf)) %>%
  arrange(well_trs_ss, year) %>%
  data.table() %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(well_trs_ss)]

#--- visual inspection ---#
# i=17
# ggplot(data=temp_match_2nd[trs==trs_ls[i],]) +
# 	geom_line(aes(y=w_to_e,x=year,color=trs_s)) +
# 	geom_point(aes(y=w_to_e,x=year,color=trs_s)) +
# 	facet_wrap(~wellid,nrow=3) +
# 	ylim(0,0.1) +
# 	theme(
# 		legend.position = 'bottom'
# 		)

# head(temp_match_2nd[trs==trs_ls[i],],50)

#++++++++++++++++
# loop over trs to find matches
#++++++++++++++++
#--- list of the trs ---#
trs_ls <- temp_match_2nd[, trs] %>% unique()
trs_len <- length(trs_ls)

matched_store_all <- list()
for (i in 1:trs_len) { # loop over trs
  print(i)
  #--- select well-energy records combinations in the temporary trs ---#
  temp_match <- temp_match_2nd[trs == trs_ls[i] & !is.na(coef_var), .(well_trs_ss, wellid, trs_s, trs_ss, coef_var)] %>%
    unique(by = "well_trs_ss")
  well_len <- temp_match[, wellid] %>%
    unique() %>%
    length()

  matched_store_i <- list()
  for (j in 1:well_len) { # loop over wells
    if (min(temp_match[, coef_var]) < 7e-4) { # if there is a match
      #--- identify the best matching well-energy combination ---#
      min_index <- which.min(temp_match[, coef_var])
      get_well <- temp_match[min_index, wellid]
      get_trs_ss <- temp_match[min_index, trs_ss]

      #--- extract the matched data and store in the temporary storage ---#
      matched_store_i[[j]] <- temp_match_2nd %>%
        filter(wellid == get_well & trs_ss == get_trs_ss) %>%
        data.table()

      #--- exclude the matched wellid and energy records  ---#
      temp_match <- temp_match %>%
        filter(!(wellid == get_well | trs_ss == get_trs_ss)) %>%
        data.table()
    } else { # if no good enough matches
      break # end the loop
    }
  }
  matched_store_all[[i]] <- rbindlist(matched_store_i)
}

SO_matched_2nd <- rbindlist(matched_store_all) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, trs_ss, district, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table() %>%
  .[, count := .N, by = wellid] %>%
  .[count > 4, ] %>%
  .[, count := NULL]

#--------------------------
# combine the 1st and 2nd matches and save
#--------------------------
SO_we <- rbind(SO_matched_1st, SO_matched_2nd) %>%
  .[, hours := kwh / (hp * 0.746)] %>%
  .[, gpm := (volaf * 12 * 27154) / (hours * 60)] %>%
  .[coef_var < 7e-4 & coef_var != 0, ] %>%
  .[w_to_e < 0.1 & w_to_e > 0.008, ]

saveRDS(SO_we, "./Data/Water_Energy/SO_energy_water.rds")

# #--- visualization ---#
# well_ls <- SO_we[,wellid] %>% unique()
# well_len <- length(well_ls)
# for (i in 1:well_len){
# 	temp_data <- SO_we[wellid==well_ls[i],]
# 	temp_cv <- temp_data[,coef_var*1e5] %>%
# 		unique() %>%
# 		round(3)
# 	ggplot(data=temp_data) +
# 		geom_line(aes(y=w_to_e,x=year)) +
# 		ylim(0,0.1) +
# 		annotate('text',label=paste('Coef Var = ',temp_cv,sep=''),y=0.075,x=temp_data[,mean(year)])
# 	ggsave(paste('./Graphs/w_to_e/Southern/consistency_',well_ls[i],'.pdf',sep=''))
# }

# ggplot(data=SO_we[w_to_e>0.008 & pwl <500,]) +
# 	geom_point(aes(y=1/w_to_e,x=pwl),size=0.4) +
# 	geom_smooth(aes(y=1/w_to_e,x=pwl)) +
# 	xlab('Pumping Water Level (feet)') +
# 	ylab('Energy Use Efficiency (kwh/acre-inch)')

# ===================================
# with Twin-Valley
# ===================================
TwinValley <- fread("./Data/Energy/EnergyProcessedByJames/TwinValley/ProcessedJames/TwinValley.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  data.tanble()

```