### 2024-08-08 Codes/1_3_proxy_quota_analysis.qmd

---
title: Welfare analysis of quota and trading system with heterogeneous production functions
---

# Objectives

This program analyzes the performance of the quota system where quota is determined randomly for a given level of target water use. This quota system imitates a quota system based on water use proxy where the measurement error exists.

+ Checks how correlation between optimal water use and measurement error affects the effectiveness of the quota system (though the answer is obvious).

# Setup

## Theoretical setup

$$
\pi = P_c \cdot f(w) - P_w \cdot w 
$$

$$
\frac{\partial\pi}{\partial w} = P_c \cdot \frac{\partial f(w)}{\partial w} - P_w 
$$

$$
\pi = P_c \cdot f(\eta \cdot p) - P_w \cdot \eta \cdot p
$$

$$
\frac{\partial \pi}{\partial p} = P_c \cdot \frac{\partial f(w)}{\partial w}\cdot \frac{\partial w}{\partial p} - P_w \cdot \eta
= \eta\cdot\frac{\partial\pi}{\partial w}
$$

What to vary:

+ variance of the measurement error 
+ direction of correlation of the measurement error and optimal water use
+ water allocation

+ positive: $\eta := \frac{w^*}{\bar{w^*}}\cdot \varepsilon$
+ negative: $\eta := \frac{\bar{w^*}}{w^*}\cdot \varepsilon$
+ independent: $\eta := \frac{\bar{w^*}}{w^*}\cdot \varepsilon$

## Prices and costs
```{r}
corn_price <- 6 # (bu/acre)
e_price <- 0.15 # ($/kwh)
w_price <- e_price * 500 / 12 # ($/inch)
other_cost <- 800 # ($/acre)
```

## Farmers yield response functions

```{r}
#--- number of farmers ---#
num_farmers <- 1000

#--- parameter heterogeneity ---#
parameters <-
  list(
    #--- ymax ---#
    # determines the maximum yield attainable
    ymax_min = 160,
    ymax_max = 250,
    #--- alpha [0, 1] ---#
    # determines rainfed yield. The higher, the lower rainfed yield.
    alpha_min = 0.4,
    alpha_max = 0.9,
    #--- beta ---#
    # determines how fast yield hits plateau as a function of water
    beta_min = 0.1,
    beta_max = 0.7
  )

#--- generate farmer data ---#
farmer_data <- gen_farmer_data_p(num_farmers, parameters)

#--- generate production data ---#
prod_data <- gen_prod_data_p(farmer_data)
```

```{r}
ggplot(farmer_data) +
  geom_histogram(aes(x = opt_w))

farmer_data[, mean(opt_w)]
```

# Simulation

Set up a parameter space to loop over.

+ `target_w`: target water use
+ `u_dev`: maximum deviation from 1 (determines the variance of measurement error)
+ `dir_cor`: direction of the correlation between (unconstrained) optimal water use and measurement error

```{r}
par_data <-
  data.table::CJ(
    target_w = seq(7, 13, by = 0.25),
    u_dev = c(0, 0.1, 0.3, 0.5),
    dir_cor = c("+", "-", "0")
  )
```

When `dir_cor = -` and `u_dev = 0`, measurement error is set so that the ratio of effective water use to proxy (`target_w`) is the same across farmers. For example, when `target_w` is the same as the average optimal water use, effective water use limit allocation maximizes social welfare (same as proxy permit trading). When `dir_cor = 0` and `u_dev = 0`, the quota system is simply a uniform quota system.

For a given value of `target_w`, `u_dev`, and `dir_cor`,

+ Create a set of quota (different quota values for different farmers) system for the number of rounds specified by the user using `make_quota_data()`
  + This function first assigns the effective water use limit and then calculate what the water to proxy conversion rate should have been
+ Conduct quota analysis using `quota_analysis_het()`

## Base case (water use quota)

The quota system with `dir_cor = 0` and `u_dev = 0` is equivalent to water quota (no measurement error).

```{r}
par_sets_w <- par_data[u_dev == 0 & dir_cor == "0", ]

results_w <- run_analysis(cases = par_sets_w, farmer_data, prod_data)
```

## Proxy quota (with measurement errors)

```{r, eval = FALSE}
par_sets_p <- par_data[!(u_dev == 0 & dir_cor == "0"), ]

results_p <- run_analysis(cases = par_sets_p, farmer_data, prod_data)

saveRDS(quota_results, "Results/quota_results_comp_stat.rds")
```

Modify the results for presentation.

```{r}
quota_results <- readRDS(here::here("Results/quota_results_comp_stat.rds"))

quota_results_het <-
  quota_results[!(dir_cor == "0" & sd == 0), ] %>%
  .[, dir_cor_txt := paste0("Correlation: ", dir_cor)] %>%
  .[, dir_cor_txt := factor(dir_cor_txt, levels = paste0("Correlation: ", c("-", "0", "+")))] %>%
  .[, sd_txt := paste0("SD: ", sd)] %>%
  .[, sd_txt := factor(sd_txt, levels = paste0("SD: ", unique(sd)))]

quota_results_no_me <-
  quota_results[dir_cor == "0" & sd == 0, ] %>%
  .[, .(w, target_w, profit)]
```

# Results Presentation

## Water use slack

Water use slack always exists when a uniform water quota is used and farmers are heterogeneous in production function.

The figure below shows water use under a range of water use limit values (x-axis) under different circumstances characterized by two parameters:

1. correlation between measurement error and unconstrained optimal water use and 
2. standard deviation of measurement error

The blue solid lines represent the water use without any measurement error. The red solid lines represent the water use with measurement error. The black solid lines represent 1 to 1 line. The vertical distance from the black line is the "slack" in water use: how much less water is used relative to the water use limit.

```{r}
ggplot() +
  geom_line(data = quota_results_het, aes(y = w, x = target_w), color = "red") +
  geom_line(data = quota_results_no_me, aes(y = w, x = target_w), color = "blue") +
  geom_abline(slope = 1) +
  facet_grid(sd_txt ~ dir_cor_txt) +
  coord_equal()
```

+ As the standard deviation of measurement error increases, water use slack becomes larger. 
+ When the correlation between measurement error and unconstrained water use is positive, water use slack with measurement error can be smaller compared to the case without measurement error. 
+ Positive correlation between them can happen in the case where energy use is used as a proxy for water use. This means that more (effective) water is allocated to those who use more water.

**Note:** For the $SD = 0$ and $Correlation = +$ case, it is programmed so that quota perfectly matches with unconstrained optimal water use.

### <span style = "color: red;"> What is the problem with water use slack? </span>

When <span style = "color: blue;"> unbiased (in estimating water use) </span> proxy is used, policy makers may unintentionally make the quota too stringent.

## Profitability

```{r}
ggplot() +
  geom_line(data = quota_results_het, aes(y = profit, x = w), color = "red") +
  geom_line(data = quota_results_no_me, aes(y = profit, x = w), color = "blue") +
  facet_grid(sd_txt ~ dir_cor_txt)
```

Given an actual water use level, farming profitability is lower with measurement error except under a small parameter space where measurement error is positively correlated with unconstrained water use and the variance of measurement error is small.

# What does water permit trading do?

Introducing trading to a water quota system increases water use if total quota stays the same (URNRD).

Reduces unfairness.

# Unfairness measure

How much loss relative to uniform water use restriction? Compare between quota and trading.