### 2024-08-08 Codes/1_5_sensitivity_to_sample_size.qmd

---
title: Examine the sensitivity of the results to the sample size of water-energy data
---

# Objective

Administrator of the quota system needs to learn the relationship between water and energy use to set the right quota to achieve water conservation target.

# Estimate energy use efficiency

```{r}
ne_data <-
  readRDS(here::here("Data/water_energy_MW.rds")) %>%
  .[nrdname != "Twin Platte", ] %>%
  .[, me := water_est - volaf] %>%
  .[, w_to_e := kwh / volaf] %>% # (kwh/feet)
  .[, region := paste0(nrdname, " (Nebraska)")] %>%
  .[w_to_e > 250, ] %>%
  .[w_to_e < 1000, ]

```

```{r}

unique_wells <- ne_data[, wellid] %>% unique()
unique_years <- ne_data[, unique(year)]

parameters <-
  data.table::CJ(
    case = 1:100,
    well_nums = length(unique_wells) * c(0.01, 0.03, 0.05, 0.1, 0.2, 0.5),
    year_nums = c(1, 3, 5, length(unique_years))
  ) %>%
  rowwise() %>%
  dplyr::mutate(sample_wells = list(
    sample(unique_wells, well_nums, replace = FALSE)
  )) %>%
  dplyr::mutate(sample_years = list(
    sample(unique_years, year_nums, replace = FALSE)
  )) %>%
  dplyr::mutate(w_to_e_estimate = list(
    ne_data[wellid %in% sample_wells, ] %>%
    .[year %in% sample_years, ] %>%
    .[, mean(w_to_e, na.rm = TRUE)]
  )) %>%
  dplyr::select(- sample_wells, - sample_years) %>%
  unnest() %>%
  data.table()
```