### 2024-08-08 Codes/2_make_figures.rmd

# Observed measurement errors

## Data Preparation

```{r}
#--- NE data ---#
ne_data <-
  readRDS(here::here("Data/water_energy_MW.rds")) %>%
  .[nrdname != "Twin Platte", ] %>%
  .[, me := water_est - volaf] %>%
  .[, w_to_e := kwh / volaf] %>% # (kwh/feet)
  .[, region := paste0(nrdname, " (Nebraska)")]

#--- CO data ---#
# Colorado data provided by Aaron Hrozencik. Pumping tests to measure kwh/acre-feet (not throughout the season like NE data).
co_data <-
  fread(here::here("Data/CO_price_data.csv")) %>%
  setnames(c("PCC", "pumpingAF"), c("w_to_e", "volaf")) %>%
  setnames(names(.), tolower(names(.))) %>%
  .[, .(wdid, year, w_to_e, volaf)] %>%
  .[, region := "Eastern Colorado"]

data <- rbind(ne_data, co_data, fill = TRUE)
```

## Distribution of energy use efficiency

```{r}
g_hist_eta <-
  ggplot(data = data[w_to_e < 1000 & w_to_e > 100, ]) +
  geom_histogram(aes(x = w_to_e), fill = NA, color = "blue", bins = 100) +
  facet_grid(region ~ .) +
  xlab("Energy use efficiency (kwh/acre-feet)") +
  ylab("Count") +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("Results/figures/g_hist_eta.pdf", g_hist_eta, height = 4, width = 6.5)
```

## Measurement error in water use estimates

```{r, fig.dim = c(6.5, 5)}
g_me <-
  ggplot(data = data[water_est < 400, ]) +
  geom_point(aes(y = volaf, x = water_est), size = 0.1) +
  facet_grid(. ~ nrdname) +
  geom_abline(intercept = 0, slope = 1) +
  xlab("Water Use Estimates Based on Energy Use (acre-feet)") +
  ylab("Observed Water Use (acre-feet)") +
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_equal() +
  scale_x_continuous(breaks = (1:9) * 100) +
  scale_y_continuous(breaks = (1:9) * 100)

ggsave("Results/figures/g_me.pdf", g_me, height = 4, width = 6.5)
```