### 2024-08-08 Codes/match_energy_water.rmd

# Preparation

```{r}
setwd("NebraskaWaterProjectsData")
```

# Import datasets

```{r}
#--------------------------
# Well registration
#--------------------------
# Notes:
# When subsec has four letters,
# The first two letters represent subsubseciton
# The last two letters represent subseciton
wells <- 
  fread("./Data/WellRegistrationRaw/Reg08_26_2016_Raw/all_rrb_wells.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  .[useid == "I", ] %>%
  .[, .(
    wellid, twnid, rngid, rngdir, section, subsecmnemonic,
    subsecid, nrdname, swl, pwl, totaldepth, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  )] %>%
  setnames(c("twnid", "rngid", "subsecmnemonic"), c("township", "range", "subsec")) %>%
  .[, `:=`(
    trs = paste(township, range, section, sep = "_"),
    subsection = ifelse(str_length(subsec) == 2, subsec, str_sub(subsec, 3, 4)),
    subsubsection = ifelse(str_length(subsec) == 2, NA, str_sub(subsec, 1, 2))
  )] %>%
  .[, trs_s := paste(trs, subsection, sep = "_")] %>%
  .[, trs_ss := paste(trs_s, subsubsection, sep = "_")] %>%
  setorder(township, range, section) %>%
  .[, N_sec := .N, by = .(trs)] %>%
  .[, N_sub := .N, by = .(trs_s)] %>%
  .[, N_subsub := .N, by = .(trs_ss)]

wells <- wells %>%
  mutate(
    dec_date = as.Date(decommissiondate, "%m/%d/%y"),
    dec_year = year(dec_date),
    dec_month = month(dec_date),
    inactive_not_dec = ifelse(is.na(dec_year) & wellstatusid == "X", 1, 0)
  ) %>%
  data.table()
```

```{r}
gw <-
  readRDS("./Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  .[, extended_id := paste(volaf, acres, year, county)] %>%
  .[, count_id := .N, by = extended_id] %>% # get rid of duplicate records
  .[count_id == 1, ]
```

## Midwest

```{r}
Midwest <- 
  readRDS("./Data/Energy/Processed/Midwest.rds") %>%
  select(year, district, rate, township, range, section, subsection, subsubsection, hp, kwh) %>%
  mutate(
    trs = paste(township, range, section, sep = "_"),
    trs_s = paste(trs, subsection, sep = "_")
  ) %>%
  arrange(township, range, section, subsection, year) %>%
  data.table() %>%
  .[, N_sec := .N, by = .(trs, year)] %>%
  .[, N_sub := .N, by = .(trs_s, year)]
#--- find wells within the Midwest PPD ---#
MW_trs_ls <- Midwest[, trs] %>% unique()
wells_MW <- wells %>%
  filter(trs %in% MW_trs_ls) %>%
  data.table()

#-----------------
# single trs-trs
#--------------------------
# pick observations that are the only record at the section level (not many)
MW_sec <- Midwest[N_sec == 1, ] %>%
  left_join(., wells_MW[N_sec == 1, ], by = "trs") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs_s = trs_s.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()
# MW_sec[!is.na(wellid),]

#--------------------------
# single trs_s-trs_s
#--------------------------
# pick observations that are the only record at the subsection level
#--- the only energy record at the subsection level ---#
MW_sub <- Midwest[N_sub == 1 & N_sec > 1, ] %>%
  left_join(., wells_MW[N_sub == 1 & N_sec > 1, ], by = "trs_s") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs = trs.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()
# matched_sub <- MW_sub[!is.na(wellid),]
# unmatched_sub <- MW_sub[is.na(wellid),]

#--------------------------
# merge with pumping data
#--------------------------
# only MW_sub and MW_sec for now. later add MW_sub_2
MW_we <- rbind(MW_sec, MW_sub) %>%
  left_join(gw, ., by = c("wellid", "year")) %>%
  filter(!is.na(district)) %>%
  data.table()
#-------------------
# pick observations that are the only record at the section level (not many)
MW_sec <- Midwest[N_sec == 1, ] %>%
  left_join(., wells_MW[N_sec == 1, ], by = "trs") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs_s = trs_s.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()
```

# Midwest
```{r}
Midwest <- 
  readRDS("./Data/Energy/Processed/Midwest.rds") %>%
  select(year, district, rate, township, range, section, subsection, subsubsection, hp, kwh) %>%
  mutate(
    trs = paste(township, range, section, sep = "_"),
    trs_s = paste(trs, subsection, sep = "_")
  ) %>%
  arrange(township, range, section, subsection, year) %>%
  data.table() %>%
  .[, N_sec := .N, by = .(trs, year)] %>%
  .[, N_sub := .N, by = .(trs_s, year)]

#--- find wells within the Midwest PPD ---#
MW_trs_ls <- Midwest[, trs] %>% unique()

wells_MW <- wells %>%
  filter(trs %in% MW_trs_ls) %>%
  data.table()

#--------------------------
# single trs-trs
#--------------------------
# pick observations that are the only record at the section level (not many)
MW_sec <- 
  Midwest[N_sec == 1, ] %>%
  left_join(., wells_MW[N_sec == 1, ], by = "trs") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs_s = trs_s.x) %>%
  dplyr::select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()
# MW_sec[!is.na(wellid),]
#--------------------------
# single trs_s-trs_s
#--------------------------
# pick observations that are the only record at the subsection level
#--- the only energy record at the subsection level ---#
MW_sub <- 
  Midwest[N_sub == 1 & N_sec > 1, ] %>%
  left_join(., wells_MW[N_sub == 1 & N_sec > 1, ], by = "trs_s") %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  rename(trs = trs.x) %>%
  dplyr::select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, district, rate, hp, kwh, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  ) %>%
  data.table()

# only MW_sub and MW_sec for now. later add MW_sub_2
MW_we <- 
  rbind(MW_sec, MW_sub) %>%
  left_join(gw, ., by = c("wellid", "year")) %>%
  filter(!is.na(district)) %>%
  data.table()
#--------------------------
MW_we <- MW_we %>%
  .[, w_to_e := volaf * 12 / kwh] %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(wellid)]

well_ls <- MW_we[, wellid] %>% unique()
well_len <- length(well_ls)

# According to visual inspections, coef_var<7e-4 seems acceptable
MW_matched_1st <- 
  MW_we[coef_var < 7e-4, ] %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

```

# multiple wells and multiple energy records in the same trs

```{r}
matched_well_ls <- MW_matched_1st[, wellid] %>% unique()
matched_trs_s_ls <- MW_matched_1st[, trs_s] %>% unique()
wells_2nd <- wells_MW[!(wellid %in% matched_well_ls), ]
e_2nd <- Midwest[!(trs_s %in% matched_trs_s_ls), ]
temp_match_2nd <- e_2nd[N_sub == 1, ] %>%
  left_join(., wells_2nd, by = "trs") %>%
  left_join(., gw, by = c("wellid", "year")) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  filter(matched == 1) %>%
  rename(trs_s = trs_s.x) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, volaf, acres, source_txt, source, hp,
    kwh, regdate, pumprate, pumpdepth, decommissiondate, district, rate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  mutate(well_trs_s = paste(wellid, trs_s, sep = "_")) %>%
  mutate(w_to_e = volaf * 12 / kwh) %>%
  filter(!is.na(volaf)) %>%
  arrange(well_trs_s, year) %>%
  data.table() %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(well_trs_s)]
#--- list of the trs ---#
trs_ls <- temp_match_2nd[, trs] %>% unique()
trs_len <- length(trs_ls)
#++++++++++++++++
matched_store_all <- list()
for (i in 1:trs_len) { # loop over trs
  #--- select well-energy records combinations in the temporary trs ---#
  temp_match <- temp_match_2nd[trs == trs_ls[i] & !is.na(coef_var), .(well_trs_s, wellid, trs_s, coef_var)] %>%
    unique(by = "well_trs_s")
  well_len <- temp_match[, wellid] %>%
    unique() %>%
    length()
  matched_store_i <- list()
  for (j in 1:well_len) { # loop over wells
    if (min(temp_match[, coef_var]) < 7e-4) { # if there is a match
      #--- identify the best matching well-energy combination ---#
      min_index <- which.min(temp_match[, coef_var])
      get_well <- temp_match[min_index, wellid]
      get_trs_s <- temp_match[min_index, trs_s]
      #--- extract the matched data and store in the temporary storage ---#
      matched_store_i[[j]] <- temp_match_2nd %>%
        filter(wellid == get_well & trs_s == get_trs_s) %>%
        data.table()
      #--- exclude the matched wellid and energy records  ---#
      temp_match <- temp_match %>%
        filter(!(wellid == get_well | trs_s == get_trs_s)) %>%
        data.table()
    } else { # if no good enough matches
      break # end the loop
    }
  }
  matched_store_all[[i]] <- rbindlist(matched_store_i)
}

MW_matched_2nd <- rbindlist(matched_store_all) %>%
  select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()

# Notes:
# According to visual inspections, coef_var<7e-4 seems acceptable

MW_matched_1st <- MW_we[coef_var < 7e-4, ] %>%
  dplyr::select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()
```

## Multiple wells and multiple energy records in the same trs

```{r}
matched_well_ls <- MW_matched_1st[, wellid] %>% unique()
matched_trs_s_ls <- MW_matched_1st[, trs_s] %>% unique()
wells_2nd <- wells_MW[!(wellid %in% matched_well_ls), ]
e_2nd <- Midwest[!(trs_s %in% matched_trs_s_ls), ]
temp_match_2nd <- e_2nd[N_sub == 1, ] %>%
  left_join(., wells_2nd, by = "trs") %>%
  left_join(., gw, by = c("wellid", "year")) %>%
  mutate(matched = ifelse(!is.na(wellid), 1, 0)) %>%
  filter(matched == 1) %>%
  rename(trs_s = trs_s.x) %>%
  dplyr::select(
    wellid, nrdname, swl, pwl, totaldepth, trs, trs_s, year, volaf, acres, source_txt, source, hp,
    kwh, regdate, pumprate, pumpdepth, decommissiondate, district, rate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd, matched
  ) %>%
  mutate(well_trs_s = paste(wellid, trs_s, sep = "_")) %>%
  mutate(w_to_e = volaf * 12 / kwh) %>%
  filter(!is.na(volaf)) %>%
  arrange(well_trs_s, year) %>%
  data.table() %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(well_trs_s)]
#--- list of the trs ---#
trs_ls <- temp_match_2nd[, trs] %>% unique()
trs_len <- length(trs_ls)
#++++++++++++++++
# loop over trs to find matches
#++++++++++++++++
matched_store_all <- list()
for (i in 1:trs_len) { # loop over trs
  #--- select well-energy records combinations in the temporary trs ---#
  temp_match <- temp_match_2nd[trs == trs_ls[i] & !is.na(coef_var), .(well_trs_s, wellid, trs_s, coef_var)] %>%
    unique(by = "well_trs_s")
  well_len <- temp_match[, wellid] %>%
    unique() %>%
    length()
  matched_store_i <- list()
  for (j in 1:well_len) { # loop over wells
    if (min(temp_match[, coef_var]) < 7e-4) { # if there is a match
      #--- identify the best matching well-energy combination ---#
      min_index <- which.min(temp_match[, coef_var])
      get_well <- temp_match[min_index, wellid]
      get_trs_s <- temp_match[min_index, trs_s]
      #--- extract the matched data and store in the temporary storage ---#
      matched_store_i[[j]] <- temp_match_2nd %>%
        filter(wellid == get_well & trs_s == get_trs_s) %>%
        data.table()
      #--- exclude the matched wellid and energy records  ---#
      temp_match <- temp_match %>%
        filter(!(wellid == get_well | trs_s == get_trs_s)) %>%
        data.table()
    } else { # if no good enough matches
      break # end the loop
    }
  }
  matched_store_all[[i]] <- rbindlist(matched_store_i)
}

MW_matched_2nd <- rbindlist(matched_store_all) %>%
  dplyr::select(
    wellid, nrdname, swl, pwl, totaldepth, year, acres, volaf, source, source_txt,
    trs, trs_s, district, rate, hp, kwh, regdate, pumprate, pumpdepth,
    decommissiondate, pumpcoldia, ownerid, wellstatusid,
    lastname, longdd, latdd, w_to_e, coef_var
  ) %>%
  data.table()
# ggplot

MW_we <- rbind(MW_matched_1st, MW_matched_2nd) %>%
  .[, hours := kwh / (hp * 0.746)] %>%
  .[, gpm := (volaf * 12 * 27154) / (hours * 60)]
MW_we[, wellid] %>% unique()
matched_all_ls <- MW_we[, wellid] %>% unique()
length(matched_all_ls)
wells_MW
unmatched_all_ls
all_wells_MW
all_wells_MW
#--------------------------
MW_we <- MW_we %>%
  .[, w_to_e := volaf * 12 / kwh] %>%
  .[is.infinite(w_to_e), w_to_e := NA] %>%
  .[, coef_var := var(w_to_e, na.rm = TRUE) / mean(w_to_e, na.rm = TRUE), by = .(wellid)]
well_ls <- MW_we[, wellid] %>% unique()
well_len <- length(well_ls)
#--- find wells within the Midwest PPD ---#
MW_trs_ls <- Midwest[, trs] %>% unique()
wells_MW <- wells %>%
  filter(trs %in% MW_trs_ls) %>%
  data.table()
wells_MW
wells_MW
matched_all_ls <- MW_we[, wellid] %>% unique()
matched_all_ls <- MW_we[, wellid] %>% unique()
u
unmatched_all_ls <- wells_MW[!(wellid %in% matched_all_ls), ]
unmatched_all_ls
matched_all_ls <- MW_we[, wellid] %>% unique()
unmatched_all_ls <- wells_MW[!(wellid %in% matched_all_ls), wellid] %>% unique()
unmatched_all_ls
matched_all_ls
length(unmatched_all_ls)
saveRDS(matched_all_ls, "./Data/Water_Energy/matched_wells_MW.rds")
saveRDS(unmatched_all_ls, "./Data/Water_Energy/unmatched_wells_MW.rds")
saveRDS(MW_trs_ls, "./Data/Water_Energy/MW_trs_ls.rds")
```

```{r}
all_gw
all_gw <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  .[, extended_id := paste(volaf, acres, year, county)] %>%
  .[, count_id := .N, by = extended_id] %>% # get rid of duplicate records
  .[count_id == 1, ]
all_gw
wells <- fread("~/Dropbox/NebraskaWaterProjectsData/Data/WellRegistrationRaw/Reg08_26_2016_Raw/all_rrb_wells.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  .[useid == "I", ] %>%
  .[, .(
    wellid, twnid, rngid, rngdir, section, subsecmnemonic,
    subsecid, nrdname, swl, pwl, totaldepth, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  )] %>%
  setnames(c("twnid", "rngid", "subsecmnemonic"), c("township", "range", "subsec")) %>%
  .[, `:=`(
    trs = paste(township, range, section, sep = "_"),
    subsection = ifelse(str_length(subsec) == 2, subsec, str_sub(subsec, 3, 4)),
    subsubsection = ifelse(str_length(subsec) == 2, NA, str_sub(subsec, 1, 2))
  )] %>%
  .[, trs_s := paste(trs, subsection, sep = "_")] %>%
  .[, trs_ss := paste(trs_s, subsubsection, sep = "_")] %>%
  setorder(township, range, section) %>%
  .[, N_sec := .N, by = .(trs)] %>%
  .[, N_sub := .N, by = .(trs_s)] %>%
  .[, N_subsub := .N, by = .(trs_ss)]
# wells[N_sub>4,]
#
wells <- fread("~/Dropbox/NebraskaWaterProjectsData/Data/WellRegistrationRaw/Reg08_26_2016_Raw/all_rrb_wells.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  .[useid == "I", ] %>%
  .[, .(
    wellid, twnid, rngid, rngdir, section, subsecmnemonic,
    subsecid, nrdname, swl, pwl, totaldepth, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  )] %>%
  setnames(c("twnid", "rngid", "subsecmnemonic"), c("township", "range", "subsec")) %>%
  .[, `:=`(
    trs = paste(township, range, section, sep = "_"),
    subsection = ifelse(str_length(subsec) == 2, subsec, str_sub(subsec, 3, 4)),
    subsubsection = ifelse(str_length(subsec) == 2, NA, str_sub(subsec, 1, 2))
  )]
all_gw <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  .[, count_id := .N, by = extended_id] %>% # get rid of duplicate records
  .[count_id == 1, ] %>%
  wells[., on = "wellid"]
all_gw <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  .[, extended_id := paste(volaf, acres, year, county)] %>%
  .[, count_id := .N, by = extended_id] %>% # get rid of duplicate records
  .[count_id == 1, ] %>%
  wells[., on = "wellid"]
all_gw <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  wells[., on = "wellid"]
all_gw
MW_wells <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  wells[., on = "wellid"] %>%
  .[trs %in% MW_trs_ls, ]
MW_wells
wells <- fread("~/Dropbox/NebraskaWaterProjectsData/Data/WellRegistrationRaw/Reg08_26_2016_Raw/all_rrb_wells.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  .[useid == "I", ] %>%
  .[, .(
    wellid, twnid, rngid, rngdir, section, subsecmnemonic,
    subsecid, nrdname, swl, pwl, totaldepth, regdate, pumprate, pumpdepth, decommissiondate,
    pumpcoldia, ownerid, wellstatusid, lastname, longdd, latdd
  )] %>%
  setnames(c("twnid", "rngid", "subsecmnemonic"), c("township", "range", "subsec")) %>%
  .[, `:=`(
    trs = paste(township, range, section, sep = "_"),
    subsection = ifelse(str_length(subsec) == 2, subsec, str_sub(subsec, 3, 4)),
    subsubsection = ifelse(str_length(subsec) == 2, NA, str_sub(subsec, 1, 2))
  )]
matched_all_ls <- readRDS("./Data/Water_Energy/matched_wells_MW.rds")
unmatched_all_ls <- readRDS("./Data/Water_Energy/unmatched_wells_MW.rds")
MW_wells <- readRDS("~/Dropbox/NebraskaWaterProjectsData/Data/Pumping/Processed/pumping.rds") %>%
  mutate(source_txt = as.character(source)) %>%
  data.table() %>%
  .[source_txt %in% c("Meter", "METER", "metered"), ] %>%
  wells[., on = "wellid"] %>%
  .[, matched := NA] %>%
  .[wellid %in% matched_all_ls, matched := TRUE] %>%
  .[wellid %in% unmatched_all_ls, matched := FALSE] %>%
  .[!is.na(matched), ]
MW_wells
MW_wells[, .(volaf, acres), by = .(matched, year)]
MW_wells[, .(volaf = mean(volaf), acres = mean(acres)), by = .(matched, year)]
MW_wells[, .(volaf = mean(volaf), acres = mean(acres)), by = .(matched)]
MW_wells[, .(volaf = mean(volaf, na.rm = TRUE), acres = mean(acres, na.rm = TRUE)), by = .(matched, year)]
MW_wells[, .(volaf = mean(volaf, na.rm = TRUE), acres = mean(acres, na.rm = TRUE)), by = .(matched)]
```