### 2024-08-08 Codes/production_function.rmd

# Data Preparation

```{r}
# Read in matlab data
yield_water <- readMat("Data/ProductionFunction_October2017.mat")

# Acess the data within the array
yiedl_water_array <- yield_water$ProductionFunction.October2017
```

# Extract and transform data by variable

Extracting crop yield and irrigation data from an array and compute expected value given an irrigation strategy

```{r}
#--------------------------
# Yield
#--------------------------
yield_data <- yiedl_water_array[[1]] %>%
  t() %>%
  data.table() %>%
  .[, smt := 1:nrow(.)] %>%
  melt(id.var = "smt") %>%
  setnames(c("variable", "value"), c("year", "yield")) %>%
  .[, yield := yield / 0.06] %>%
  .[, year := as.numeric(gsub("V", "", year))] %>%
  setkey(smt, year)

#--- Expected crop yield ---#
e_yield <- yield_data[, .(e_yield = mean(yield)), by = smt] %>%
  setkey(smt)

#--------------------------
# Irrigation
#--------------------------
irr_data <- yiedl_water_array[[2]] %>%
  t() %>%
  data.table() %>%
  .[, smt := 1:nrow(.)] %>%
  melt(id.var = "smt") %>%
  setnames(c("variable", "value"), c("year", "ir")) %>%
  .[, year := as.numeric(gsub("V", "", year))] %>%
  .[, ir := ir * 0.0393701] %>%
  setkey(smt, year)

#--- expected irrigation ---#
e_ir <- irr_data[, .(e_ir = mean(ir)), by = smt] %>%
  setkey(smt)

#--------------------------
# Combine and save
#--------------------------
#--- raw data ---#
raw_data <- yield_data[irr_data]
```

# Find the maximum yield by irrigation bin for each year

```{r}
year_ls <- raw_data[, year] %>% unique()
prod_store <- list()

for (y in year_ls) {
  temp_data <- 
    copy(raw_data[year == y, ]) %>%
    .[, ir_bin := cut(ir, c(quantile(ir, prob = seq(0, 1, by = 0.05))) %>% unique())] %>%
    .[, ir_bin := as.character(ir_bin)] %>%
    .[is.na(ir_bin), ir_bin := "0"] %>%
    .[, .(yield = max(yield), ir = mean(ir)), by = ir_bin] %>%
    .[, ir_bin := NULL] %>%
    #--- get rid of observations that saw a decline in yield  ---#
    setorder(ir) %>%
    .[, yield_prev := shift(yield, n = 1, type = "lag")] %>%
    .[, dif := yield - yield_prev] %>%
    .[dif > 0 | is.na(dif), ] %>%
    .[, .(yield, ir)]

  if (y == 11) { # special treatment for year 12
    temp_data[ir == 0, yield := 194]
  }
  # g_reg <- ggplot(data=reg_data) +
  # 	geom_point(aes(y=yield,x=ir))
  # ggsave(g_reg,file=paste('./Figures/reg_data_y_',y,'.pdf',sep=''))

  #--- augment data for a better fitting ---#
  max_yield <- temp_data[, max(yield)]
  max_ir <- temp_data[, max(ir)]
  min_ir_data <- temp_data[ir == min(ir), ][rep(1, 5), ]
  max_ir_data <- data.table(
    yield = rep(max_yield, 15),
    ir = max_ir + seq(0, max_ir * 0.2, length = 15)
  )
  reg_data <- rbind(min_ir_data, temp_data, max_ir_data)
  prod_store[[as.character(y)]] <- scam(yield ~ s(ir, k = 7, bs = "micv"), data = reg_data)
}

saveRDS(prod_store, "Results/prod_fcn.rds")
```

# Visualize the production functions

```{r}
for (y in year_ls) {
  temp_prod <- prod_store[[paste(y)]]
  g_temp <- data.table(ir = seq(
    floor(min(temp_prod$model$ir) * 100) / 100,
    floor(max(temp_prod$model$ir) * 100) / 100,
    by = 0.05
  )) %>%
    .[, y_hat := predict(temp_prod, newdata = .)] %>%
    ggplot(data = .) +
    geom_line(aes(y = y_hat, x = ir))
  ggsave(g_temp, file = paste("Figures/production_function/prod_y_", y, ".pdf", sep = ""))
}
```