### 2024-08-08 Writing/paper_structure.rmd

# Main

# Results

## ME in water use estiamtes based on energy

## Quota

Energy quota relative to water quota

+ uniform allocation for both energy and water

Things to look at:

+ Economic loss (gain) using 
+ Water use slack

Comparative statics wrt

+ the variance of the measurement error
+ heterogeneity of production function

## Trading

+ Trading reduces profit loss
+ Reduces uncertainty in water use

Measurement error is absorbed.

## Incentive to improve energy use efficiency

+ Energy use looks the same! Trouble for the administrator.
+ But, water use has gone up

## Fairness

# Method

## Data

## Simulation

+ Heterogeneous farmers

# Supplementary Appendix

## More general results (other cases)